MS&E 226: Project Part 1 Code

### 2024-12-03 MS&E 226_Project Part 2_Code.Rmd
2c and 2d with OLS fit from lasso coefficients

#### (c)
```{r 2c Fit to test data OLS}

@@

@@

# Extract coefficients from the test set
ols_test_coefs <- coef(ols_test)

# Extract coefficients from training and holdout models
training_coefs <- as.matrix(coef(ols_train))
holdout_coefs <- as.matrix(ols_test_coefs)

@@

@@

>>>>

#### (c)

```{r 2c Fit to test data OLS}

@@

# Save a copy of the holdout dataset to be transformed/cleaned 
# Create df_test with only the specified columns from holdout
df_test <- holdout[, column_names]

@@

# Extract p-values from the training and test models
training_pvalues <- summary(ols_train)$coefficients[, "Pr(>|t|)"]
test_pvalues <- summary(ols_test)$coefficients[, "Pr(>|t|)"]

# Remove the intercept term from p-values
training_pvalues <- training_pvalues[names(training_pvalues) != "(Intercept)"]
test_pvalues <- test_pvalues[names(test_pvalues) != "(Intercept)"]

# Ensure feature names match between the training and holdout p-values
common_features <- intersect(names(training_pvalues), names(test_pvalues))

# Subset p-values to only include common features
training_pvalues <- training_pvalues[common_features]
test_pvalues <- test_pvalues[common_features]

# Combine p-values into a comparison data frame
comparison <- data.frame(
  Predictor = common_features,
  Training_Pvalues = round(training_pvalues, 8),
  Test_Pvalues = round(test_pvalues, 8)
)

# Sort the comparison data frame in increasing order of Training_Pvalues
comparison <- comparison[order(comparison$Training_Pvalues), ]

# Display the comparison
print("Comparison of p-values from training and test data:")
print(comparison)

## compare coefficients

@@

# Function to fit OLS on bootstrap sample and return coefficients
bootstrap_ols <- function(data, indices) {
  # Bootstrap sampling
  boot_sample <- data[indices, ]

  # Fit OLS model
  ols_fit <- lm(ols_formula, data = boot_sample)

  # Extract coefficients
  coefs <- coef(ols_fit)

  # Return coefficients (excluding intercept if not needed)
  return(coefs)
}

# Set up bootstrap
set.seed(123)
n_boot <- 1000  # Number of bootstrap samples

# Perform bootstrap
bootstrap_results <- boot(
  data = df_pred,
  statistic = bootstrap_ols,
  R = n_boot
)

@@

# Format as a data frame
cis_df <- data.frame(
  Predictor = names(bootstrap_results$t0),
  CI_Lower = bootstrap_cis[1, ],
  CI_Upper = bootstrap_cis[2, ]
)

# Extract OLS confidence intervals from the standard regression output
ols_cis_standardized <- confint(ols_test, level = 0.95)
ols_cis_df_standardized <- data.frame(
  Predictor = rownames(ols_cis_standardized),
  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
)

# Merge with Bootstrap confidence intervals for comparison
comparison_ols <- merge(
  cis_df,  # Bootstrap confidence intervals from OLS
  ols_cis_df_standardized,
  by = "Predictor",
  all.x = TRUE
)

# Display the comparison
print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
print(comparison_ols)

# Visualize the difference in Bootstrap vs. Standardized OLS CIs

# Calculate confidence interval sizes for Bootstrap and OLS
comparison_ols$Bootstrap_CI_Size <- comparison_ols$CI_Upper - comparison_ols$CI_Lower
comparison_ols$OLS_CI_Size <- comparison_ols$OLS_CI_Upper_Standardized - comparison_ols$OLS_CI_Lower_Standardized

# Reshape the data for plotting
ci_data_ols <- comparison_ols %>%
  select(Predictor, Bootstrap_CI_Size, OLS_CI_Size) %>%
  pivot_longer(cols = c(Bootstrap_CI_Size, OLS_CI_Size), 
               names_to = "Method", 
               values_to = "CI_Size")

# Rename methods for better readability
ci_data_ols$Method <- ifelse(ci_data_ols$Method == "Bootstrap_CI_Size", "Bootstrap OLS", "Standard OLS Output")

# Create the density plot
ggplot(ci_data_ols, aes(x = CI_Size, fill = Method)) +
  geom_density(alpha = 0.5) +
  scale_x_log10() +
  labs(
    title = "Density Plot of Confidence Interval Sizes: Bootstrap vs. Standard OLS Output",
    x = "Confidence Interval Size",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    legend.title = element_blank()
  )

# Violin plot
ggplot(ci_data_ols, aes(x = Method, y = CI_Size)) +
  geom_violin(fill = "lightblue") +
  geom_boxplot(width = 0.1, fill = "white") +
  theme_minimal() +
  labs(title = "Distribution of CI_Size by Group", x = "Method", y = "CI_Size")
